### 2015-02-10 paper/system_architecture.tex

\section{System Design}
In this section we describe key aspects of our system design which increase the throughput of inter node communication. First of all it is worth to note that even with one TCP connection, it is possible to saturate the network link usage if the sending and receiving processes send and receive byte arrays without a delay. However serializing java object messages to byte arrays and deserializing back takes considerable amount of time. Therefore we can improve the communication performance by reducing this message processing time. We use multitasking and bulk message serializing to solve this problem.

In this design, we use thread pools both at client side and server side to processes messages while sharing a single TCP connection. Firstly both thread pools access underline connection resource only to send receive byte arrays making message serialization and deserialization processes parallel. Secondly it serializes messages together to improve serialization process.  Figure \ref{interprocess} shows the design involved in sending a message from one process to another.

First the client process thread sends this message to its \textit{ElementContainer}. \textit{ElementContainer} holds all the streams (this is an abstraction of a link in the process graph) to which this message needs to be send and it pushes the message to all the streams. Stream decides the target node and buffers the messages for a configured amount of messages. Then it passes the buffered  message list along the target node details to \textit{ConnectionManager} which holds \textit{ClientConnection}s for each node. \textit{ConnectionManager} serializes the message list and sends the byte array to correct \textit{ClientConnection} using the target node. \textit{ClientConnection} uses the \textit{DataOutput} which wraps the TCP connection to send message to server side.

At the server side there is a set of \textit{ServerTask}s which reads receiving messages from a pool of \textit{DataInput} streams available in the \textit{ServerConnection} (we register these connections at the connections creating time) and send them to receiving processes. First a server task acquires a \textit{DataInput} stream and reads the binary message and release it. Then it deserializes the binary message using the sending process id to identify the event type.  After that it passes this message to the \textit{WorkerContainer} which holds all processes. Finally \textit{WorkerContainer} dispatches this message to correct process using receiving process id. Message ordering happens at the process level if required.

\begin{figure}[!tii]
        \centering
        \includegraphics[width=3.0in]{interprocess.png}
        \caption{Communication between two Process}
        \label{interprocess}
\end{figure}